detect patches

### 2024-02-13 scripts/classify_patches.Rmd
r to vector

landscapR <- function(x, pout, pRout){

  taxon <- gsub('[.]tif', '', basename(x))
  x <- terra::rast(file.path('../results/patches', x))
  x <- terra::ifel(x > 0, 1, NA)
  x <- terra::mask(x, terra::crop(hydr_bound, x), inverse = TRUE)
  landscape <- landscapemetrics::get_patches(x, directions = 4)
  calcs <- landscapemetrics::calculate_lsm(x, what = c('lsm_c_enn_mn', 'lsm_c_enn_cv',
                "lsm_p_enn", "lsm_p_cai", "lsm_p_para", "lsm_p_frac"), 
              neighbourhood = 4, directions = 4)

  p <- '../results/'
  write.csv(
    calcs[calcs$level == 'patch', c('id', 'metric', 'value')],
    file = paste0(p, pout, '/', taxon, '-', 'patch_vals.csv'),  row.names = F
  )

  write.csv(
    calcs[calcs$level == 'class', c('metric', 'value')],
    file = paste0(p, pout, '/', taxon, '-', 'class_vals.csv'), row.names = F
  )

  writeRaster(landscape[["layer_1"]][["class_1"]], 
    filename = paste0(p, pRout, '/', taxon, '-', 'patchIDS.tif')
  )

@@

combine_pop_patches <- function(x){

  pIDs <- read.csv(paste0('../results/occupiedPatches/', x, '-topSuitability.csv'))
  top_patches <- terra::rast(paste0('../results/fine_patches/', x, '-patchIDS.tif'))
  highest_no <- minmax(top_patches)[2]
  top_patches <- terra::subst(top_patches, pIDs$PatchID,  pIDs$PatchID, others = NA)
  names(top_patches)<- 'patches'
  pts <- st_read(paste0('../data/raw/occurrence/combined_records/', x, '.shp'), quiet = TRUE) |>
    dplyr::filter(Occurrence == 1)
  marg_patch <- terra::rast(paste0('../results/marginal_habitat/', x, '.tif'))

  fsize <- file.info(paste0('../results/marginal_habitat/', x, '.tif'))$size/1e9
  if(fsize < 0.025){marg_patch <- terra::patches(marg_patch, directions = 4)} else
    if(fsize > 0.025){marge1 <- terra::patches(terra::aggregate(marg_patch, 2, fun = 'mean'), directions = 4)} else
      if(fsize > 0.04){marge1 <- terra::patches(terra::aggregate(marg_patch, 3, fun = 'mean'), directions = 4)} else
        if(fsize > 0.05){marge1 <- terra::patches(terra::aggregate(marg_patch, 4, fun = 'mean'), directions = 4)} else
          if(fsize > 0.06){marge1 <- terra::patches(terra::aggregate(marg_patch, 5, fun = 'mean'), directions = 4)} else
            if(fsize > 0.07){marge1 <- terra::patches(terra::aggregate(marg_patch, 6, fun = 'mean'), directions = 4)}
  if(exists('marge1')){marg_patch <- terra::resample(marge1, marg_patch, threads = 16, method = 'near')}

  occ_pts <- terra::extract(marg_patch, terra::vect(pts), bind = TRUE) |>
    sf::st_as_sf()
  occ_poly <- dplyr::filter(occ_pts, is.na(patches)) |>
    sf::st_buffer(250) # we cut out 90, but let's go a bit further for corner cases.
  occ_poly <- terra::extract(marg_patch, terra::vect(occ_poly), weights = TRUE) %>% 
    dplyr::group_by(ID) |> # now make sure we get the patch with the 'most' overlap with the polygon
    tidyr::drop_na(patches) |> # these values actually NA. 
    dplyr::slice_max(weight, with_ties = F)

  occ_patches <- c(
    dplyr::filter(occ_pts, ! is.na(patches)) |>
      dplyr::pull(patches), 
    occ_poly$patches
  )

  if(length(occ_patches) > 0){

    marg_patch <- terra::subst(marg_patch, occ_patches,  occ_patches, others = NA)
    old_ID <- unique(marg_patch)$patches # gather and relabel patches 
    lkp <- data.frame(
      old_ID = old_ID,
      new_ID = highest_no + 1:length(old_ID)
    )
    marg_patch <- terra::subst(marg_patch, lkp$old_ID, lkp$new_ID)
    final_patches <- terra::cover(top_patches, marg_patch)

    terra::writeRaster(final_patches, paste0('../results/allOccupiedPatches/', x, '.tif'))
  } else {
    terra::writeRaster(top_patches, paste0('../results/allOccupiedPatches/', x, '.tif'))
  }
}

@@

```{r create a polygon dataset, eval = F}

@@

>>>>

@@

@@

```{r create polygon datasets of patches, eval = F}

@@

f <- paste0('../results/fine_patches/', 
        list.files('../results/fine_patches/'))

  sf::st_write(pg, paste0('../results/fine_patchesVector/', taxon, '.shp'), quiet = TRUE)

}
lapply(f, tif2shp)
